---
title: "Wikpedia Rockband Timeline Figure for nf-core/mag"
author: "James A. Fellows Yates"
format: html
editor: visual
---

This notebook describes how you can generate a [Wikipedia band-member timeline plot](https://en.wikipedia.org/wiki/List_of_Iron_Maiden_band_members#Timeline) from a Git repo.

## Get raw data

To get raw data, you can use the following command on a Git repo (make sure to update the location of repo)

We do a shallow clone to make the cloning process faster, generate the contribution information from the `log` command, and then delete the shallow clone after.

```{bash}
git clone --depth 1 git@github.com:nf-core/mag.git
git -C ~/git/nf-core/mag --no-pager log --pretty=format:"%h%x09%an%x09%ad%x09%s" --date=iso | cut -f 2-3 > "nf-core-mag_contribs.tsv"
rm -rf mag/
```

## Clean raw data

Load relevant libraries

```{r}
library(readr)
library(janitor)
library(tidyr)
library(stringr)
library(dplyr)
library(lubridate)
library(tibble)
library(ggplot2)
library(paletteer)

sessionInfo()

```

## Data cleaning and summarising

We can then load the data and do some basic cleaning.

```{r}
data_raw <- read_tsv(paste0("nf-core-mag_contribs.tsv"), col_names = c("committer", "date_of_commit")) |> clean_names()

data_committers <- data_raw |>
  select(committer) |>
  unique() |>
  arrange(committer) |>
  write_tsv(paste0("nf-core-mag_contribs_list.tsv"))
```

Now we have to do some manual clean up of the git log output, as in some cases we may only have GitHub handle(s!) rather than a name, and also we want to annotate with organisation each member was/is a member of.

I have separately made a user-name-org database file manually that we can load and bind onto our git log data

```{r}
data_contributor_db <- read_tsv("nf-core-mag-contribs_db.tsv")

data_full <- left_join(data_raw, data_contributor_db)
```

We can then use this information plus other information to filter and summarise the table with the commit history, namely: remove bots, and remove users with less than a given number of commits.

In this case we will only include contributors with more than 10 commits.

```{r}
data_filtered_n_commits <- data_full |> 
  group_by(name) |> 
  summarise(n_commits = n()) |> 
  arrange(desc(n_commits))  |> 
  filter(name != "bot", n_commits >= 10)
```

We can then filter the table for the people with more than 10 commits, and then calculate and then calculate their weekly commits.

```{r}
data_weekly_commits <- data_full |>
  mutate(week = cut.Date(as.Date(date_of_commit), breaks = "1 week")) |> 
  select(name, organisation, organisation_colour, week) |> 
  filter(name %in% data_filtered_n_commits$name)
```

We can also extract the timestamp for the plot

```{r}
data_timestamp <- data_full |> select(date_of_commit) |> arrange(date_of_commit) |> last() |> pull() |> as_date()
```

Now we can reduce our full history, and then extract their first and last commit date. We will then reorder and fix the order of names for plotting purposes.

```{r}
data_final_start_stop <- data_weekly_commits |>
  select(name, week, organisation) |>
  filter(name %in% data_filtered_n_commits$name) |>
  group_by(name) |>
  summarise(
    first_commit = as_date(last(week)),
    latest_commit = as_date(first(week)),
    organisation = unique(organisation)
  ) |>
  arrange(first_commit, latest_commit) |>
  mutate(name = factor(name, levels = name),
         organisation = factor(organisation, levels = unique(organisation)),
         first_commit = first_commit - 30,
         latest_commit = latest_commit + 30
         )

```

## Plotting

Here is a basic plot using a palette with a sufficient number of colours

```{r}
plot_wikipedia_style <- ggplot(data_final_start_stop, aes(x = first_commit, y = name, colour = organisation)) +
 geom_segment(aes(xend = latest_commit, yend = name, linewidth = 10)) +
 geom_point(data = data_weekly_commits, aes(x = as_date(week), y = name), colour = 'white', size = 0.3) +
 scale_x_date(date_breaks = "1 year", date_minor_breaks = "4 months", date_labels = "%Y", labels = label_at(1)) +
 scale_y_discrete(limits = rev(levels(data_final_start_stop$name))) +
 theme_classic(base_family = 'Inter') +
 scale_colour_paletteer_d("LaCroixColoR::paired", ) +
 theme(legend.position = "bottom", text = element_text(family = "Inter")) +
 guides(linewidth = "none") +
 xlab("Date") +
 ylab("Contributor (with 10 or more commits)") +
 labs(colour = 'Organisation')

plot_wikipedia_style
```

To make a nicer looking image we can take the primary colours of each organisation and use this to fill the bars.

This information I included in the source file for `data_contributor_db`, and colour picked from each organisations website.

```{r}
palettes_orgs <- data_contributor_db |> 
  select(organisation, organisation_colour) |> 
  distinct() |> 
  filter(!is.na(organisation)) |> 
  deframe()
```

Finally we can try to plot the timeline

```{r}
plot_wikipedia_style_orgcolours <- plot_wikipedia_style + scale_colour_manual(values = palettes_orgs)

plot_wikipedia_style_orgcolours

```

We can also customise this slightly to indicate handovers of primary development teams

```{r, fig.width=8, height=2}
theme_alternating_grid <- function() {
  theme_classic() +
    theme(
      panel.grid.major.y = element_line(color = "grey90", size = 3.7),
      panel.grid.minor.y = element_line(color = "white"),
      text = element_text(family = "Inter", size = 8)
    )
}

plot_wikipedia_style_orgcoloursannotations <- plot_wikipedia_style_orgcolours +
  theme_alternating_grid() +
  geom_vline(xintercept = as.numeric(date('2019-12-21'))) + ## Hadrien - Sabrina/Daniel handover, based on first release made on GH
  geom_vline(xintercept = as.numeric(date('2022-06-14'))) +  ## Sabrina/Daniel - James handover, based on first release made on GH 
  geom_segment(data = data_final_start_stop |> filter(name == 'Hadrien Gourlé'), xend = as.numeric(date('2019-12-21')),  aes(yend = name, linewidth = 5), colour = '#FFFFFF') +
  geom_segment(data = data_final_start_stop |> filter(name == 'Sabrina Krakau'), xend = as.numeric(date('2022-06-14')), aes(yend = name, linewidth = 5), colour = '#FFFFFF') +
  geom_segment(data = data_final_start_stop |> filter(name == 'Daniel Straub'),  x = as.numeric(date('2019-12-21')), xend = as.numeric(date('2022-06-14')), aes(yend = name, linewidth = 5), colour = '#FFFFFF') +
  geom_segment(data = data_final_start_stop |> filter(name == 'James A. Fellows Yates'), x=as.numeric(date('2022-06-14')), aes(xend = latest_commit, linewidth = 5), colour = '#FFFFFF') +
labs(caption = paste0('As of ', as.character(data_timestamp)))


plot_wikipedia_style_orgcoloursannotations

```

```{r}
ggsave(plot_wikipedia_style_orgcoloursannotations, filename = 'nf-core-mag-contributors-timeline-plot.png', device = 'png', height=107, width = 178, units = 'mm', dpi=300)
ggsave(plot_wikipedia_style_orgcoloursannotations, filename = 'nf-core-mag-contributors-timeline-plot.svg', device = 'svg', height=107, width = 178, units = 'mm', dpi=300)

```
